### 2012-01-25 article/Article.tex
refonte biblio, MAJ des affiliations, et ajout de figures

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

\usepackage{color}

% to be able to use options in graphics
\usepackage{graphicx}
% for pseudo code
\usepackage{listings}
% subfigures
\usepackage{subfigure}

% some maths symbols
\usepackage{amstext,amssymb}

% to include accents in bliblio
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  hyperref should be the last package to be loaded.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}

 \definecolor{lightgray}{gray}{.75}
\newcommand{\commentaire}[1]{
\colorbox{lightgray}{#1}
}

%  This is a template for Papers to the Insight Journal. 
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{Efficient N-Dimensional surface estimation using Crofton formula and run-length encoding}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
% \release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann$^{1,2}$ {\small{and}} David Legland$^{3,4,5}$}

\authoraddress{\small{
{$^1$}INRA, UMR 1198 Biologie du D\'eveloppement et Reproduction, Jouy-en-Josas, F-78350, France\\
{$^2$}ENVA, UMR 1198 Biologie du D\'eveloppement et Reproduction, Jouy-en-Josas, F-78350, France\\
{$^3$}INRA, UMR 0782 G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France\\
{$^4$}AgroParisTech, UMR 0782 G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France\\
{$^5$}INRA, UMR 1318 Institut Jean-Pierre Bourgin, Versailles, F-78026, France
}}

\begin{document}
\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

\begin{abstract}
\noindent
Unlike the measure of the area in 2D or of the volume in 3D, the perimeter and the surface are not easily measurable in a discretized image.
In this article we describe a method based on the Crofton formula to measure those two parameters in a discritized image. The accuracy of
the method is discussed and tested on several known objects. An algorithm based on the run-length encoding of binary objects is presented
and compared to a brute force approach.
An implementation is provided and integrated in the LabelObject/LabelMap framework contributed earlier by the authors.
\end{abstract}

\tableofcontents

\section{Introduction}

Surface area and perimeter are widely used parameters describing the size of objects observed in images, 
and are commonly used for computing various shape factors.
Unlike the area that can be easily measured by counting pixels belonging to the object, measuring the 
perimeter is not as straightforward, 
and naive methods can lead to huge systematic errors \cite{Klette2004, Legland2007}.
Usually, the countour of the object is extracted, then its length is measured.
The same approach is usually applied to 3D images: the boundary of the 3D object is first reconstructed,
e.g. by using marching cubes \cite{Lorensen1987}, then its surface area is computed by summing 
area of individual triangles.

The Crofton method is an alternative method that allows measuring the perimeter of 2D objects,
the surface area of 3D objects, and more generally the $(d-1)$-dimensional measure of $d$-dimensional 
object boundary. 
\commentaire{Ajouter un petit argumentaire sur convergence des mesures avec Crofton.}

The article first recalls the mathematical principles of $d$-perimeter measures in digital images
using the Crofton method. An efficient implementation based on run-length encoding is then presented.
The method is evaluated on various $2D$ and $3D$ synthetic shapes and compared with other methods.
The effect on the roundness shape factor is also investigated.

\section{Principles}
\subsection{Crofton formula}

% Crofton formula
Perimeter measure of 2D objects, surface area measure of 3D objects 
and more generally $(d-1)$-surface measure of $d$-dimensional objects 
can be expressed in a unified formalism by using the Crofton formula.
This formula consists in integrating the intercept number of the object with
lines of various orientation and positions. Its expression in
the general form is:
\begin{eqnarray}
S^{(d-1)}(X) & = & \frac{d \cdot v_d}{\pi}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
\end{eqnarray}
where $X$ is the structure of interest, $v_d$ is the volume of the $d$-dimensional ball,
$\mathcal{L}^d$ is the set of all lines in the $d$-dimensional space. 
The above integral is normalised such that the mass of lines hitting the unit ball 
equals $v_{d-1}$, the measure of the unit ball projection on a $(d-1)$-dimensional plane.

The Euler-Poincar\'e characteristic $\chi$ is equal to number of connected components
of the intersection of $X$ with a line $L$, or equivalently half the number of intersections
of the boundary of $X$ with the line.

% 2D and 3D
For planar and 3D cases, Equation 1 can be rewritten:
\begin{eqnarray}
P(X) & = & \pi \int_{\mathcal{L}^2} \chi \left( X \cap L \right) dL \\
S(X) & = &  4  \int_{\mathcal{L}^3} \chi \left( X \cap L \right) dL
\end{eqnarray}

\subsection{Crofton formula in discrete images}

The Crofton formula can be easily applied to discrete binary images \cite{Lang2001, Legland2007}. 
The integral over lines can be decomposed into an integral over a finite set of directions
and an integral over lines parallel to a given direction.
The number of connected components in the intersections of the structure with a discrete line
is obtained by counting the number of intercepts with its boundary.
\commentaire{ 
Ajouter une formule avec des sommes a la place des integrales (David)}

\commentaire{dire comment on calcule les poids (1) des espaces entre les droites et (2) des poids associes aux directions}

\commentaire{Cas des directions principales ($d$ en $nd$) et de toutes les directions (4 en 2D et 13 en 3D)}

\subsection{Intercepts count in digital images}

The intercepts count, required to use the method described earlier, can be efficiently computed by using the run-length
encoding of the binary objects.

\subsection{Roundness}

The roundness is defined the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
The equivalent $d$-perimeter is defined as the perimeter of the $d$-ball with same volume $V$.
\begin{eqnarray}
roundness             & = & \frac{P_{eq}}{P} \\
P_{eq}                & = & \frac{d \cdot V_{eq}}{R_{eq}} \\
R_{eq}                & = & \sqrt[\displaystyle d]{\frac{V\cdot\Gamma(\frac{d+1}{2})}{\pi^\frac{d}{2}}} \\
\Gamma(\frac{d+1}{2}) & = & \begin{cases}
                               \displaystyle \frac{d}{2}!  &\text{if}~ d ~\text{is even} \\
                               \displaystyle \frac{\sqrt{\pi} \cdot d!!}{2^{\frac{d+1}{2}}}  &\text{if}~ d ~\text{is odd} \\
                            \end{cases} \\
d!!                   & = & \begin{cases}
                                1 &\text{if}~ d \leq 1 \\
                                d \cdot (d-2)!!  & \text{otherwise} \\
                             \end{cases}
\end{eqnarray}

Roundness takes values between $0$ and $1$. 
It is equal to $1$ for circular or spherical shapes, 
and decreases for more complicated shapes.

\section{Implementation}

The implementation of this algorithm has been done in the \verb$itk::ShapeLabelMapFilter$ class, which is already in charge of the computation of
several shape descriptors. The run-length encoding used in the \verb$itk::LabelObject$ representation is reused.
The implementation is N-dimensional and thus is usable for any image dimension. The most useful cases, 2D and 3D have been specialized
to provide a more accurate estimation by also using the diagonals -- 4 directions in 2D and 13 directions in 3D. In the 4D case and greater,
the diagonals are not used.

\subsection{N-Dimensional name}
The attribute is called {\em Perimeter} independently of the dimension of the image and is available in the \verb$itk::ShapeLabelObject$.

\subsection{Border management}

The borders of the image is considered to be in the background, so the contour of an object touching the border of the image is measured.
This was not the case in the previous (undocumented) implementation and has been proven to be misleading for many users.
It is possible to subtract the perimeter on the border to the full perimeter if the measure without the part touching the border is needed.

\subsection{Multithreading}

The architecture implemented in \verb$itk::ShapeLabelMapFilter$ use one thread per LabelObject. The perimeter estimation has been integrated in this
architecture, providing a multithreaded implementation as long as there are several LabelObjects in the input LabelMap.

The perimeter estimation is now turned on by default, but can still disabled with \verb$SetComputePerimeter(false)$ in
\verb$itk::ShapeLabelMapFilter$ if not needed. This should enhance the user experience, especially for the attributes non obviously derived
from the perimeter like the roundness.

\section{Evaluation method}

\subsection{Test shapes}

Perimeter and surface area are measured on discretization of various 2D and 3D shapes 
whose true perimeter or surface area is known.

Planar test shapes include disks, rings obtained by the difference of several disks,
trefoil shape, ellipses with $a=XX$ and $b=XX$, and rectangle with length $XX$ and width $XX$.

Test shapes for 3D measurements include
balls with various radii, hollow ball, prolate and oblate ellipsoids, and cuboids. 
For each type of shape, several orientations were considered.

\subsection{Shape discretization}

A 2D or 3D binary image can be considered as a subset of a rectangular grid $L_d$ where
d = 2,3. Such a grid can be written as
\begin{eqnarray}
L_d = \Delta_{1} \mathbb{Z} \times \ldots \times \Delta_{1} \mathbb{Z},
\end{eqnarray}

where the $d$-uple $(\Delta_{1},\ldots,\Delta_{d})$ defines the pixel
or voxel size. A digitization scheme represents how a continuous
structure $X$ is digitized into a binary image $X_\text{d}$. 
According to the
covering scheme a grid point $x$ belongs to $X_\text{d}$ if the grid
cell centered on $x$ hits $X$ (See Fig.~\ref{fig:DiskDiscretization}).

% Example of reconstruction of a discrete disk
\begin{figure}[!htb]
\begin{center}
\includegraphics[width=5cm]{images/discreteDisk}
\end{center}
\caption{Discretization of a disk}
\label{fig:DiskDiscretization}
\end{figure}

\subsection{Measurements}

For each test shape, measurements were made on
several discretisation of the shape, with various orientations and various origins.

The center of each shape was chosen at random within the center pixel or voxel of the grid.

Perimeter of 2D discretized shapes was measured using Crofton method using 2 and 4 directions, 
and by computing the length of the reconstructed contour (Matlab Software was used).
Surface area of 3D discretized shapes was measured using Crofton method using 3 and 13 directions, 
and by computing the surface area of the reconstruced shape obtained by the marching cubes algorithm.
The average and the standard deviation of measurements were computed for all shapes discretized
with the same resolution. Effect of orientation was investigated by computing average and standard 
deviation of shapes having the same orientation.

\subsection{Timing evaluation}

???

\section{Accuracy}

\subsection{Perimeter of disks at various resolutions}

Figure \ref{fig:MeasureDiskPerimeter} shows average and standard deviation of perimeter
measured on disks with various resolutions. Perimeter measured using Crofton method converges
towards the real value when the resolution increase. 
The convergence speed is faster when four directions are used instead of two, 
and the variability is lower.

% Example of reconstruction of a discrete disk
\begin{figure}[!htb]
\begin{center}
\includegraphics[width=10cm]{images/perimeterDisks}
\end{center}
\caption{Perimeter measure of disks with various resolutions and at random positions}
\label{fig:MeasureDiskPerimeter}
\end{figure}

Perimeters measured with Matlab show similar variability, but the measured values do not
converge toward the true value. This shows the inaccuracy of the method consisting in
counting boundary pixels.

The oscillations of the variability is due to the fact that for disks with diameters equal to an
integer, the number of intersections with horizontal or vertical lines is always the same, 
thus greatly reducing the variability of measures.

\subsection{Perimeter of squares with various orientations}

Figure \ref{fig:MeasureSquareOriented} shows measures obtained on discretized squares (side 30 pixels)
with various orientations. All methods oscillate around the expected value. 
Using Crofton method with only 2 directions can produce greater errors on the measure than with 4 
directions.

% Example of reconstruction of a discrete squares with various orientations
\begin{figure}[!htb]
\begin{center}
\includegraphics[width=8cm]{images/errorSquareRotated}
\includegraphics[width=8cm]{images/errorRelSquareRotated}
\end{center}
\caption{Variations of perimeter measure of a digitized square depending on the orientation}
\label{fig:MeasureSquareOriented}
\end{figure}

\section{Timing}

[c'est rapide]

\section{Application}

\subsection{ICOPAN}

Methods were applied on images acquired within the ICOPAN project \cite{Andrey2010}.

\appendix

\bibliographystyle{plain}
\bibliography{InsightJournal}
\nocite{ITKSoftwareGuide}

\end{document}